String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-18 2023-1-19/sections/cefa.tex
all review finished

In the sequel, we define three CEFA operations: union, intersection, and concatenation. The following section will use these three operations to solve RECL constraints. \denghang{Note that the union, intersection, and concatenation operations are defined in a slightly more involved manner than symbolic register automata~\cite{sra} and counter automata~\cite{GGM12}, as a result of the technical choice that the accepting conditions of CEFA do not distinguish final states.}

>>>>

In the sequel, we define three CEFA operations: union, intersection, and concatenation. The following section will use these three operations to solve RECL constraints. \denghang{Note that the union, intersection, and concatenation operations are defined in a slightly more involved manner than symbolic register automata~\cite{sra} and counter automata~\cite{GGM12}, as a result of the technical choice that the accepting condition is attached to each automaton, not each final state.}

### 2023-10-18 2023-1-19/sections/implementation.tex
all review finished

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. \denghang{The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Are our decision procedure more efficient than the state-of-the-art string constraint solvers on RECL constraints? (Sec \ref{subsec:overall_eval})
  \item [Q2:] Why our decision procedure is more efficient?(Sec \ref{subsec:large_bounds_eval})
  \item [Q3:] Are the size-reduction techniques practical?(Sec \ref{subsec:size_reduction_eval})
  \item [Q4:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker? (Sec \ref{subsec:size_reduction_eval})
\end{itemize}
}

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. \denghang{Other industrial benchmarks suites are not utilized because they contains no counting operator.} There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

@@

% \medskip
% \noindent
% \emph{Overall evaluation.} 
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, \denghang{and the results distinct to the ground truth are called \emph{soundness errors} in the result table.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). \denghang{To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.}

>>>>

%!TEX root = ../main.tex
%\documentclass{standalone}
%\begin{document}
In this section, we list our experimental results. We first present the overall experiment results on RegCoL and AutomatArk benchmark suites. Then we present experiment results of the large-counting part and evaluate the simplification techniques. \denghang{The goal of our experiments is to answer the following questions:
\begin{itemize}
  \item [Q1:] Are our decision procedure more efficient than the state-of-the-art string constraint solvers on RECL constraints? (Sec \ref{subsec:overall_eval})
  \item [Q2:] Why is our decision procedure more efficient? (Sec \ref{subsec:large_bounds_eval})
  \item [Q3:] Are the size-reduction techniques practical? (Sec \ref{subsec:size_reduction_eval})
  \item [Q4:] How effective is the algorithm for solving the nonemptiness problem compared to the case that uses the nuXmv model checker~\cite{atva2020}? (Sec \ref{subsec:size_reduction_eval})
\end{itemize}
}

@@

Our experiments utilize two benchmark suites, namely, \emph{RegCoL} and \emph{AutomatArk}. \denghang{Other industrial benchmarks suites are not utilized because they contains no counting operators.} There are 48,843 instances in total, and all benchmark instances are in the SMTLIB2 format.
Moreover, it turns out that only 5\% of regexes among the 48,843 instances are non-register-representable (see Section~\ref{subsec:regex2cefa}).

@@

% \medskip
% \noindent
% \emph{Overall evaluation.} 
The experiment results can be found in Table~\ref{tab:results_regcol}. Note that we take the result of CVC5 as the ground truth\footnote{Initially,  we used the majority vote of the results of the solvers as the ground truth. Nevertheless, on some problem instances, all the results of the three solvers in the Z3 family are wrong (after manual inspection), thus failing this approach on these instances.}, \denghang{and the results distinct to the ground truth are called \emph{soundness errors}.} We can see that $\ostrichrecl$ solves almost all 48,843 instances, except 182 of them, that is, it solves \textbf{48,662} instances correctly. The number is %3,994/2,272/24,577/12,832/1,091/7,092 
\textbf{3,908/1,111/12,838/2,306/2,396 more} than the number of instances solved by CVC5/Z3str3RE/Z3str3/Z3seq/OSTRICH respectively.
%      
%A soundness error is reported if the result of a solver for an instance is different from CVC5. 
% their results on some instances are inconsistent with the results of the other three solvers, i.e. CVC5, OSTRICH, and OSTRICH$^{\rm RECL}$.
%
Moreover, $\ostrichrecl$ is the second fastest solver, whose average time on each instance is close to the fastest solver Z3str3RE (\textbf{1.93s}/1.62s). \denghang{To answer \textbf{Q1}, \textbf{$\ostrichrecl$ is the second fastest solver on these benchmark suites, and it solves the most number of instances}.}

### 2023-10-18 2023-1-19/sections/introduction.tex
all review finished

\medskip
\noindent
\emph{Related work.} 
We discuss more related work about regexes with counting in programming languages.  
The determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.
Real-world regexes in programming languages include features beyond classical regular expressions, e.g., the greedy/lazy Kleene star, capturing groups, and back references. \denghang{Symbolic register automata\cite{sra} are raised to handle back references.} Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22} \denghang{and \cite{sra}}.  Solving string constraints involving lengths has also been investigated by various researchers, see e.g. \cite{WC+18,atva2020,z3str3re}. \denghang{Rule-based string solvers like Z3str/2/3/4/sequences and CVC4/5 are incapable of handling regexes with counting and complementary, even for the simple constraint $x\in [0-8]^{\{1,200\}}\wedge x\not\in [0-9]^+$. Automata-based string solvers like Z3Str3RE and Ostrich can solve it, but they are inefficient because they have to build automata by unwinding counting operator, resulting in an automaton with more than $1800$ states for regex $[0-8]^{\{1,200\}}$. Moreover, the counting operator is not supported by string solvers Z3-Trau and Slent.}

>>>>

\medskip
\noindent
\emph{Related work.} 
We discuss more related work about regexes with counting in programming languages.  
The determinism of regexes with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata were proposed in~\cite{redos_lenka,HS+23} to quickly match a subclass of regexes with counting. Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regexes with counting.   Nevertheless, the nonemptiness problem of these automata was not considered, and it is unclear whether these automata models can be used for solving string constraints with regex-counting and string-length.
Real-world regexes in programming languages include features beyond classical regular expressions, e.g., the greedy/lazy Kleene star, capturing groups, and back references. \denghang{Symbolic register automata\cite{sra} are raised to handle back references.} Real-world regexes have been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving~\cite{CF+22}. Nevertheless, the counting operators are still unfolded explicitly in~\cite{CF+22} \denghang{and \cite{sra}}.  Solving string constraints involving lengths has also been investigated by various researchers, see e.g. \cite{WC+18,atva2020,z3str3re}. \denghang{Rule-based string solvers like Z3str/2/3/4/sequences and CVC4/5 are incapable of handling regexes with counting and complementary, even for the simple constraint $x\in [0-8]^{\{1,200\}}\wedge x\not\in [0-9]^+$. Automata-based string solvers like Z3Str3RE and Ostrich can solve it, but they are inefficient because they have to build automata by unwinding counting operator, resulting in state explosion problem. Moreover, the counting operator is not supported by string solvers Z3-Trau and Slent.}
